The LHC as a Neutrino-Ion Collider

### 2023-11-09 journal_sub/sec-dis-pseudodata.tex
some referee updates
https://docs.google.com/document/d/1Uaoc4B43W9kNWLhwCEfg5po59jjxrRj-qubZol22BsY/edit?usp=sharing

Different neutrino structure functions provide complementary sensitivity
 to the partonic flavour decompositions of nucleons.
 %
 To illustrate this feature, consider a leading order  calculation
 for a proton target with $n_f=4$ active quark flavours,
a diagonal CKM matrix, and no heavy quark mass effects.
 %
 The resulting $F_2^{\nu p}$ and $xF_3^{\nu p}$ structure functions read
\begin{align}
	 F_2^{\nu p}(x,Q^2) &= 2x\lp f_{\bar{u}} + f_{d} + f_{s} + f_{\bar{c}} \rp(x,Q^2) \, , \nonumber  \\
	F_2^{\bar{\nu} p}(x,Q^2) &= 2x\lp f_u + f_{\bar{d}} + f_{\bar{s}} + f_c \rp(x,Q^2) \, , \label{eq:neutrinoSFs_proton} \\
	xF_3^{\nu p}(x,Q^2) &= 2x\lp -f_{\bar{u}} + f_d +f_s - f_{\bar{c}}\rp(x,Q^2)  \, , \nonumber\\
	xF_3^{\bar{\nu} p}(x,Q^2) &= 2x\lp f_u - f_{\bar{d}} -f_{\bar{s}} + f_{c}\rp(x,Q^2) \, . \nonumber
\end{align}
 The corresponding expressions for a neutron target are obtained from isospin symmetry
 \begin{align}
 	 F_2^{\nu n}(x,Q^2) &= 2x\lp f_{\bar{d}} + f_{u} + f_{s} + f_{\bar{c}} \rp(x,Q^2) \, , \nonumber  \\
 	F_2^{\bar{\nu} n}(x,Q^2) &= 2x\lp f_d + f_{\bar{u}} + f_{\bar{s}} + f_c \rp(x,Q^2) \, , \label{eq:antineutrinoSFs_neutron} \\
 	xF_3^{\nu n}(x,Q^2) &= 2x\lp -f_{\bar{d}} + f_u +f_s - f_{\bar{c}}\rp(x,Q^2)  \, , \nonumber\\
 	xF_3^{\bar{\nu} n}(x,Q^2) &= 2x\lp f_d - f_{\bar{u}} -f_{\bar{s}} + f_{c}\rp(x,Q^2) \, , \nonumber
 \end{align}
 while for an isoscalar, free-nucleon target denoted by $N$ one has
 \begin{align}
 	\noindent
 	&F_2^{\nu N}(x,Q^2)= 2x\lp f_{u^+} + f_{d^+} + 2f_s + 2f_{\bar{c}} \rp(x,Q^2), \nonumber  \\
 	&F_2^{\bar{\nu} N}(x,Q^2)= 2x\lp f_{u^+} + f_{d^+} + 2f_{\bar{s}} + 2f_c \rp(x,Q^2), \label{eq:neutrinoSFs_isoscalar} \\
 	&xF_3^{\nu N}(x,Q^2)= 2x\lp f_{u^-} + f_{d^-} +2f_s - 2f_{\bar{c}}\rp(x,Q^2), \nonumber\\
 	&xF_3^{\bar{\nu} N}(x,Q^2)= 2x\lp   f_{u^-} + f_{d^-}-2f_{\bar{s}} +2 f_{c}\rp(x,Q^2), \nonumber \qquad
 \end{align}
 in terms of the valence and sea PDF combinations defined by the following relations
\begin{align}
	 f_{q^+} (x,Q^2) &\equiv \lp f_{q}+f_{\bar{q}}\rp(x,Q^2) \, , \nonumber \\
	f_{q^-} (x,Q^2) &\equiv \lp f_{q}- f_{\bar{q}}\rp(x,Q^2) \, .
\end{align}
 We note that, even for isoscalar targets, separate measurements
 for neutrinos and antineutrinos will not be equivalent, since in general
 the strange and charm sea asymmetries $f_{s^-}$ and
 $f_{c^-}$ are not expected to vanish.

@@

\paragraph{FASER$\nu$.}
%
The ForwArd Search ExpeRiment (FASER) detector
and its companion FASER$\nu$~\cite{FASER:2019aik,FASER:2019dxq,FASER:2023zcr,FASER:2022hcn}
are located at the TI12 tunnel of the CERN accelerator complex.
%
Both detectors are aligned
with the collision axis line-of-sight (LOS)
and have been acquiring data since the beginning of Run III in 2022.
%
Neutrino scattering takes place in the FASER$\nu$ 
detector, composed by interleaved emulsion and tungsten plates and
adding up to a mass of 1.1 tonnes with a fiducial volume of $20~\rm{cm} \times 25~\rm{cm} \times 80~{\rm cm}$.
%
The FASER apparatus is immersed in a magnetic field,  providing charged-lepton
identification thanks to two 1 m-long dipole magnets with $B=0.57$ T
and another 1.5 m-long magnet in front of the spectrometer. 
%
Neutrino detection and identification can be carried out either using the emulsion
films, which have the key benefit of excellent position and angular resolution,
or instead using the electronic detector components of FASER, which enable the tagging
of the outgoing downstream energetic muons.
%
FASER$\nu$ is sensitive to neutrinos with pseudorapidity $\eta_\nu \ge 8.5$
and can also identify charm-tagged events.
%
To identify the charge of the lepton from a neutrino interaction, the lepton is required to pass through the FASER spectrometer.
%
This imposes an angular requirement on the outgoing charged lepton, indicated
in Table~\ref{tab:FPF_experiments}.
%
Also, to identify a DIS interaction the emulsion detector requires at least 5 charged tracks
to emerge from the interaction vertex.
%
We implement this last constraint
in our simulations by requiring a minimum hadronic energy when calculating the event yields.

@@

 \paragraph{AdvSND.}
 %
 This proposed experiment consists actually on  two detectors, a far-detector to be installed
 at the FPF with a coverage in neutrino pseudorapidity of $7.2 \le \eta_\nu \le 8.4$
 (i.e., off-LOS, same as SND@LHC) and a near detector installed somewhere else in the LHC
 complex and covering the range $4 \le \eta_\nu \le 5$.
 %
Here we focus on the former.
 %
 It would be
 composed (from upstream to downstream) by a target region
 for  vertex reconstruction and electromagnetic energy measurement, followed  by a hadronic calorimeter, a  muon 
 identification system, and finally  a magnet enabling muon charge and momentum measurements.
 %
 The target region of the detector, where the neutrino interactions take place, is made of thin sensitive layers 
 interleaved with tungsten plates, for a total mass of 5 tonnes and fiducial length of 50 cm.
 %
 This detector configuration will be able to track muons with energy $E_\nu \gsim 20$ GeV
 within an acceptance of 100 mrad and provide information on the charge
 of the  outgoing muon thanks to its magnet.
 %
 The total energy of the hadronic final state will be measured
 in the hadronic calorimeter.
 %
 No information on the expected performance of the AdvSND-far detector
 is available and hence no estimate of the systematic errors
 is carried out.

 \paragraph{FLArE.}
 %
 Building upon recent progress in liquid noble gas neutrino detectors over the last decade (ICARUS, MicroBooNE, SBND, DUNE), this experiment
 would rely on a modularized liquid argon (LAr) time projection detector.
 %
 The use of LAr as a target is beneficial for final-state particle identification, track angle, and kinetic energy measurements with sub-millimeter spatial resolution in all dimensions.
 %
 The detector will be equipped with a magnetized hadron/muon calorimeter downstream of the liquid argon volume
 for muon charge and momentum measurements.
 %
 While muon neutrinos with energy $E_\nu \gsim 2$ GeV would not be fully
 contained in the  FLArE detector,
for our projections
we assume that outgoing high-energy
muons would be measured in the FASER2 spectrometer
  situated downstream in the FPF cavern.

>>>>

Different neutrino structure functions provide complementary sensitivity
 to the partonic flavour decompositions of nucleons.
 %
 To illustrate this feature, consider a leading order  calculation
 for a proton target with $n_f=4$ active quark flavours,
a diagonal CKM matrix, and no heavy quark mass effects.
 %
 The resulting $F_2^{\nu p}$ and $xF_3^{\nu p}$ structure functions read
\begin{align}
	 F_2^{\nu p}(x,Q^2) &= 2x\lp f_{\bar{u}} + f_{d} + f_{s} + f_{\bar{c}} \rp(x,Q^2) \, , \nonumber  \\
	F_2^{\bar{\nu} p}(x,Q^2) &= 2x\lp f_u + f_{\bar{d}} + f_{\bar{s}} + f_c \rp(x,Q^2) \, , \label{eq:neutrinoSFs_proton} \\
	xF_3^{\nu p}(x,Q^2) &= 2x\lp -f_{\bar{u}} + f_d +f_s - f_{\bar{c}}\rp(x,Q^2)  \, , \nonumber\\
	xF_3^{\bar{\nu} p}(x,Q^2) &= 2x\lp f_u - f_{\bar{d}} -f_{\bar{s}} + f_{c}\rp(x,Q^2) \, . \nonumber
\end{align}
 The corresponding expressions for a neutron target are obtained from isospin symmetry
 \begin{align}
 	 F_2^{\nu n}(x,Q^2) &= 2x\lp f_{\bar{d}} + f_{u} + f_{s} + f_{\bar{c}} \rp(x,Q^2) \, , \nonumber  \\
 	F_2^{\bar{\nu} n}(x,Q^2) &= 2x\lp f_d + f_{\bar{u}} + f_{\bar{s}} + f_c \rp(x,Q^2) \, , \label{eq:antineutrinoSFs_neutron} \\
 	xF_3^{\nu n}(x,Q^2) &= 2x\lp -f_{\bar{d}} + f_u +f_s - f_{\bar{c}}\rp(x,Q^2)  \, , \nonumber\\
 	xF_3^{\bar{\nu} n}(x,Q^2) &= 2x\lp f_d - f_{\bar{u}} -f_{\bar{s}} + f_{c}\rp(x,Q^2) \, , \nonumber
 \end{align}
 while for an isoscalar, free-nucleon target denoted by $N$ one has
 \begin{align}
 	\noindent
 	&F_2^{\nu N}(x,Q^2)= 2x\lp f_{u^+} + f_{d^+} + 2f_s + 2f_{\bar{c}} \rp(x,Q^2), \nonumber  \\
 	&F_2^{\bar{\nu} N}(x,Q^2)= 2x\lp f_{u^+} + f_{d^+} + 2f_{\bar{s}} + 2f_c \rp(x,Q^2), \label{eq:neutrinoSFs_isoscalar} \\
 	&xF_3^{\nu N}(x,Q^2)= 2x\lp f_{u^-} + f_{d^-} +2f_s - 2f_{\bar{c}}\rp(x,Q^2), \nonumber\\
 	&xF_3^{\bar{\nu} N}(x,Q^2)= 2x\lp   f_{u^-} + f_{d^-}-2f_{\bar{s}} +2 f_{c}\rp(x,Q^2), \nonumber \qquad
 \end{align}
 in terms of the valence and sea PDF combinations defined by the following relations
\begin{align}
	 f_{q^+} (x,Q^2) &\equiv \lp f_{q}+f_{\bar{q}}\rp(x,Q^2) \, , \nonumber \\
	f_{q^-} (x,Q^2) &\equiv \lp f_{q}- f_{\bar{q}}\rp(x,Q^2) \, .
\end{align}
 We note that, even for isoscalar targets, separate measurements
 for neutrinos and antineutrinos will not be equivalent, since in general
 the strange and charm sea asymmetries $f_{s^-}$ and
 $f_{c^-}$ are not expected to vanish~\cite{Sufian:2018cpj,Sufian:2020coz}.

@@

\paragraph{FASER$\nu$.}
%
The ForwArd Search ExpeRiment (FASER) detector
and its companion FASER$\nu$~\cite{FASER:2019aik,FASER:2019dxq,FASER:2023zcr,FASER:2022hcn}
are located at the TI12 tunnel of the CERN accelerator complex.
%
Both detectors are aligned
with the collision axis line-of-sight (LOS)
and have been acquiring data since the beginning of Run III in 2022.
%
Neutrino scattering takes place in the FASER$\nu$ 
detector, composed by interleaved emulsion and tungsten plates and
adding up to a mass of 1.1 tonnes with a fiducial volume of $20~\rm{cm} \times 25~\rm{cm} \times 80~{\rm cm}$.
%
The FASER apparatus is immersed in a magnetic field,  providing charged-lepton
identification thanks to two 1 m-long dipole magnets with $B=0.57$ T
and another 1.5 m-long magnet in front of the spectrometer. 
%
Neutrino detection and identification can be carried out either using the emulsion
films, which have the key benefit of excellent position and angular resolution,
or instead using the electronic detector components of FASER, which enable the tagging
of the outgoing downstream energetic muons.
%
FASER$\nu$ is sensitive to neutrinos with pseudorapidity $\eta_\nu \ge 8.5$
and can also identify charm-tagged events.
%
To identify the charge of the lepton from a neutrino interaction, the lepton is required to pass through the FASER spectrometer.
%
This imposes an angular requirement on the outgoing charged lepton, indicated
in Table~\ref{tab:FPF_experiments}.
%
Also, to identify a DIS interaction the emulsion detector requires at least 5 charged tracks
to emerge from the interaction vertex.
%
We implement this last constraint
in our simulations by requiring a minimum hadronic energy when calculating the event yields, as the charged track multiplicity is expected to grow with $W$~\cite{Aachen-Bonn-CERN-Munich-Oxford:1981lfk,FASER:2019dxq}.

@@

 \paragraph{AdvSND.}
 %
 This proposed experiment~\cite{Feng:2022inv} consists actually on  two detectors, a far-detector to be installed
 at the FPF with a coverage in neutrino pseudorapidity of $7.2 \le \eta_\nu \le 8.4$
 (i.e., off-LOS, same as SND@LHC) and a near detector installed somewhere else in the LHC
 complex and covering the range $4 \le \eta_\nu \le 5$.
 %
Here we focus on the former.
 %
 It would be
 composed (from upstream to downstream) by a target region
 for  vertex reconstruction and electromagnetic energy measurement, followed  by a hadronic calorimeter, a  muon 
 identification system, and finally  a magnet enabling muon charge and momentum measurements.
 %
 The target region of the detector, where the neutrino interactions take place, is made of thin sensitive layers 
 interleaved with tungsten plates, for a total mass of 5 tonnes and fiducial length of 50 cm.
 %
 This detector configuration will be able to track muons with energy $E_\nu \gsim 20$ GeV
 within an acceptance of 100 mrad and provide information on the charge
 of the  outgoing muon thanks to its magnet.
 %
 The total energy of the hadronic final state will be measured
 in the hadronic calorimeter.
 %
 No information on the expected performance of the AdvSND-far detector
 is available and hence no estimate of the systematic errors
 is carried out.

 \paragraph{FLArE.}
 %
 Building upon recent progress in liquid noble gas neutrino detectors over the last decade (ICARUS, MicroBooNE, SBND, DUNE), this experiment~\cite{Batell:2021blf,Feng:2022inv}
 would rely on a modularized liquid argon (LAr) time projection detector.
 %
 The use of LAr as a target is beneficial for final-state particle identification, track angle, and kinetic energy measurements with sub-millimeter spatial resolution in all dimensions.
 %
 The detector will be equipped with a magnetized hadron/muon calorimeter downstream of the liquid argon volume
 for muon charge and momentum measurements.
 %
 While muon neutrinos with energy $E_\nu \gsim 2$ GeV would not be fully
 contained in the  FLArE detector,
for our projections
we assume that outgoing high-energy
muons would be measured in the FASER2 spectrometer
  situated downstream in the FPF cavern.
